---
title: "Homework02"
format: html
editor: visual
---

# **Task 1 : Basic Vector Practice**

## 1. Create 2 vectors

```{r}
Pre_blood_pressure = c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
post_blood_pressure = c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

## 2.Assign names to the vector elements using the paste() function.

```{r}
#Note that names() can be overwritten by a character vector Pre_vector
subject_data <- paste("Subject", 1:20, sep = "_")
#Patient = subject_data
names(Pre_blood_pressure) <- subject_data
names(post_blood_pressure) <- subject_data

```

## 3. Calculate the change in blood pressure for each patient by subtracting post-treatment measurements from pre-treatment measurements.

```{r}
#Recall that R does math element-wise!
decrease_blood_pressure <- Pre_blood_pressure - post_blood_pressure 
cat("Change in blood pressure for each patient:\n", decrease_blood_pressure)

```

## 4. Calculate the average decrease in blood pressure across all patients.

```{r}
avg_blood_pressure <- mean(decrease_blood_pressure)
cat("The average Blood pressure across all patients: ", avg_blood_pressure)
```

## 5. Determine which patients experienced a decrease in blood pressure after treatment (a positive change).

## Use the which() function to just return the indices (and names) associated with this type of change.

```{r}
which(decrease_blood_pressure > 0)      
cat("Number of Patients who decrease in blood Pressure are:-",length(which(decrease_blood_pressure > 0)))
```

## 6. Subset the vector of differences to only return those that have a positive change.

```{r}
# A Boolean used as an indexing vector returns only indices were a TRUE occurred.)
# it show only True values
subset_dec_BP <- decrease_blood_pressure[decrease_blood_pressure > 0]
subset_dec_BP
```

## It will show True and False both

```{r}

subset_dec_true_bp <- decrease_blood_pressure > 0
subset_dec_true_bp
```

## 7. Calculate the average decrease in blood pressure for those where the blood pressure decreased (positive change).

```{r}
cat("Average decrease in Blood Pressure :- ",mean(subset_dec_BP))
```

# **Task 2: Basic Data Frame practice**

## 1. Create a data frame object with four columns corresponding to your data above: patient, pre_bp, post_bp, and diff_bp

```{r}
#Patient_data = data.frame(subject_data, Pre_blood_pressure, post_blood_pressure, decrease_blood_pressure)
bp_df = data.frame(pre_bp = Pre_blood_pressure, 
                           post_bp = post_blood_pressure, 
                           diff_bp = decrease_blood_pressure)
bp_df

#bp_df <- bp_df[!duplicated(bp_df), ]
#bp_df
```

## 2.Return only rows of the data frame where the diff_bp column is negative

```{r}
#(Use [ or learn about the subset() function if you’d like. If you use [, don’t reference the original vector from the first part, access the column of the data frame to make your comparison with 0.)

neg_diff_bp <- subset(bp_df, diff_bp < 0 )
neg_diff_bp


```

## 3. Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120 and FALSE if it is not.

### Create a new column based on condition

```{r}
bp_df$new_post_bp <- ifelse(bp_df$post_bp < 120, "True", "False")
bp_df

```

## 4. print the data frame knitr::kable(bp_df)

```{r}
#print("Final Patient BP data", knitr::kable(bp_df))
#print(knitr::kable(bp_df))
knitr::kable(bp_df, caption = "**Final Patient BP Data**")

```

# **Task 3:**

## 1. We have the following pre-treatment values for subjects 1 through 10 (different set of subjects):

### • 138, 135, 147, 117, 152, 134, 114, 121, 131, 130

## treatment, the subjects were measured again (subjects 1 through 10 match)

### • 105, 136, 123, 130, 134, 143, 135, 139, 120, 124

## Create a vector with pre and post bp data

```{r}
control_pre_bp <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
control_post_bp <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)


```

## Give the name to treatment pre and post data

```{r}
control_subject <- paste("Subject", 1:10, sep = "_")
names(control_pre_bp) <- control_subject
names(control_post_bp) <- control_subject


```

## Calculate the change in each patient

```{r}
control_diff_bp <- control_pre_bp - control_post_bp 
cat("Change in blood pressure for each patient:\n", control_diff_bp)


```

## Calculate the average decrease in blood pressure across all patients.

```{r}
control_avg_bp <- mean(control_diff_bp)
cat("The average Blood pressure across all patients: ", control_avg_bp)

```

## Determine which patients experienced a decrease in blood pressure after treatment

```{r}
which(control_diff_bp > 0)      
cat("Number of Patients who decrease in blood Pressure are:-",length(which(control_diff_bp > 0)))

```

## Subset the vector of differences to only return those that have a positive change.

```{r}
control_subset_dec_BP <- control_diff_bp[control_diff_bp > 0]
control_subset_dec_BP


```

## Calculate the average decrease in blood pressure for those where the blood pressure decreased

```{r}

control_avg_dec_bp <- mean(control_subset_dec_BP)
control_avg_dec_bp

```

## Create a Data Frame

```{r}
control_bp_df = data.frame(control_pre_bp, 
                           control_post_bp, 
                           control_diff_bp)
control_bp_df

```

## Return only rows of the data frame where the diff_bp column is negative

```{r}
control_neg_diff_bp <- subset(bp_df, diff_bp < 0 )
control_neg_diff_bp

```

## Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120 and FALSE if it is not.

```{r}
control_bp_df$new_treat_post_bp <- ifelse(control_bp_df$control_post_bp < 120, "True", "False")
control_bp_df



```

## Final data frame

```{r}
knitr::kable(control_bp_df, caption = "**Final Patient Treatment BP Data**")


```

## 2. Now create and store a list with two elements:

### • 1st element named treatment and contains the first data frame you created.

### • 2nd element named placebo and contains the second data frame you created.

```{r}

Final_data <- list(
  treatment = bp_df,
  placebo = control_bp_df
)

Final_data

str(Final_data)
```

## 3. Access the first list element using three different types of syntax (the result can be a list or just the data frame).

```{r}
print("**First syntax to show the result of first data frame**")
Final_data$treatment

print("** second syntax to show the result of first data frame**")
Final_data[[1]]

print("** Third syntax to show the result of first data frame**")
getElement(Final_data, "treatment")



```

## 4. Use your list from question 2. In one line of code, access the pre_bp column of the placebo data frame.

```{r}

print("Access the pre_bp column\n")
Final_data[[2]][1]
```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```
